---
title: "Farver og striber"
keypoints:
- Use the `dplyr` package to manipulate dataframes.

objectives:
- Describe the purpose of an R package and the **`dplyr`** and **`tidyr`** packages.

questions:
- How can I select specific rows and/or columns from a dataframe?
teaching: 50
exercises: 30
source: Rmd
---


```{r, include = FALSE}
source("../bin/chunk-options.R")
knitr_fig_path("03-")
source("../bin/download_data.R")
```


## Color blind friendly colors ##


> ## Note
>
> The packages in the tidyverse, namely **`dplyr`**, **`tidyr`** and **`ggplot2`**
> accept both the British (e.g. *summarise*) and American (e.g. *summarize*) spelling
> variants of different function and option names. For this lesson, we utilize
> the American spellings of different functions; however, feel free to use
> the regional variant for where you are teaching.
{: .callout}


efterfølgende ting knækker hvis vi piller ved denne kodeblok, så vi beholder
den til vi har fjernet disse efterfølgende ting.

```{r, results = 'hide', purl = FALSE, message = FALSE}

## load the tidyverse
library(tidyverse)

interviews <- read_csv("data/SAFI_clean.csv", na = "NULL")

## inspect the data
interviews

## preview the data
# view(interviews)
```




```{r, purl = FALSE}
interviews2 <- filter(interviews, village == "Chirodzo")
interviews_ch <- select(interviews2, village:respondent_wall_type)
```

```{r, purl = FALSE}
interviews_ch <- select(filter(interviews, village == "Chirodzo"),
                         village:respondent_wall_type)
```


```{r, purl = FALSE}
interviews_ch <- interviews %>%
    filter(village == "Chirodzo") %>%
    select(village:respondent_wall_type)

interviews_ch

```


> ## Exercise
>
>  Using pipes, subset the `interviews` data to include interviews
> where respondents were members of an irrigation association
> (`memb_assoc`) and retain only the columns `affect_conflicts`,
> `liv_count`, and `no_meals`.
>
> > ## Solution
> >
> > ```{r}
> > interviews %>%
> >     filter(memb_assoc == "yes") %>%
> >     select(affect_conflicts, liv_count, no_meals)
> > ```
> {: .solution}
{: .challenge}







```{r, purl = FALSE}
interviews_long <- interviews_wide %>%
    pivot_longer(cols = c(burntbricks, cement, muddaub, sunbricks),
                 names_to = "respondent_wall_type",
                 values_to = "wall_type_logical") %>%
    filter(wall_type_logical) %>%
    select(-wall_type_logical)
```


```{r, purl = FALSE}
interviews_items_owned <- interviews %>%
  separate_rows(items_owned, sep = ";") %>%
  replace_na(list(items_owned = "no_listed_items")) %>%
  mutate(items_owned_logical = TRUE) %>%
    pivot_wider(names_from = items_owned,
                values_from = items_owned_logical,
                values_fill = list(items_owned_logical = FALSE))

nrow(interviews_items_owned)
```


```{r, eval = FALSE}
interviews_items_owned <- interviews %>%
```


```{r, eval = FALSE}
replace_na(list(items_owned = "no_listed_items")) %>%
```


```{r, eval = FALSE}
mutate(items_owned_logical = TRUE) %>%
```

```{r, eval = FALSE}
pivot_wider(names_from = items_owned,
            values_from = items_owned_logical,
            values_fill = list(items_owned_logical = FALSE))

```

View the `interviews_items_owned` dataframe. It should have
`r nrow(interviews)` rows (the same number of rows you had originally), but
extra columns for each item. How many columns were added?






```{r, purl = FALSE}
interviews_plotting <- interviews %>%
  ## pivot wider by items_owned
  separate_rows(items_owned, sep = ";") %>%
  ## if there were no items listed, changing NA to no_listed_items
  replace_na(list(items_owned = "no_listed_items")) %>%
  mutate(items_owned_logical = TRUE) %>%
  pivot_wider(names_from = items_owned,
              values_from = items_owned_logical,
              values_fill = list(items_owned_logical = FALSE)) %>%
  ## pivot wider by months_lack_food
  separate_rows(months_lack_food, sep = ";") %>%
  mutate(months_lack_food_logical = TRUE) %>%
  pivot_wider(names_from = months_lack_food,
              values_from = months_lack_food_logical,
              values_fill = list(months_lack_food_logical = FALSE)) %>%
  ## add some summary columns
  mutate(number_months_lack_food = rowSums(select(., Jan:May))) %>%
  mutate(number_items = rowSums(select(., bicycle:car)))
```

Now we can save this dataframe to our `data_output` directory.

```{r, purl = FALSE, eval = FALSE}
write_csv (interviews_plotting, file = "data_output/interviews_plotting.csv")
```

```{r, purl = FALSE, eval = TRUE, echo = FALSE}
if (!dir.exists("data_output")) dir.create("data_output")
write_csv(interviews_plotting, "data_output/interviews_plotting.csv")
```

{% include links.md %}
