---
title: "Processing JSON data (Optional)"
teaching: 30
exercises: 15
questions:
- "What is JSON format?"

objectives:
- "Describe the JSON data format"
keypoints:
- "JSON is a popular data format for transferring data used by a great many Web based APIs"

source: Rmd
---

```{r, include=FALSE}
source("../bin/chunk-options.R")
knitr_fig_path("05-")
source("../bin/download_data.R")
## load the tidyverse
library(tidyverse)
```

# gode til fordelinger

## violin

## density

## Histogram

## Boxplot

## Ridgeline

# Gode til korrelationer

## scatter
Basic scatterplot
```{r, purl = FALSE, eval = TRUE, echo = FALSE}
ggplot(iris, aes(x=Sepal.Length, y=Sepal.Width)) + 
    geom_point()
```

alternativ:
```{r, purl = FALSE, eval = TRUE, echo = FALSE}
ggplot(iris, aes(x=Sepal.Length, y=Sepal.Width)) + 
    geom_point(
        color="black",
        fill="#69b3a2",
        shape=22,
        alpha=0.5,
        size=6,
        stroke = 1
        ) 
```


## Heatmap

## Correlogram

## Bubble

## Connected scatter

## Density 2d

# Gode til ranking

## Barplot


```{r, purl = FALSE, eval = TRUE, echo = FALSE}
ggplot(mtcars, aes(x=as.factor(cyl), fill=as.factor(cyl) )) +  
  geom_bar( ) +
  scale_fill_manual(values = c("red", "green", "blue") ) +
  theme(legend.position="none")
```
If there is no natural order of the x-values, consider reordering 
by size:

```{r, purl = FALSE, eval = TRUE, echo = FALSE}
mtcars %>% 
  mutate(cyl = factor(cyl)) %>% 
  mutate(cyl = fct_infreq(cyl)) %>% 
ggplot(aes(x=cyl, fill=as.factor(cyl) )) +  
  geom_bar( ) +
  scale_fill_manual(values = c("red", "green", "blue") ) +
  theme(legend.position="none")
```


## Spider/radarplot

## Wordcloud

## Parallel

## Lollipop

## Cirkulær barplot

# gode til at vise dele af en helhed

## grupperet og stakket barplot

## Treemap

## Doughnut

## Pie Chart

## Dendrogram

## Circular packing

# Good for showing evolution

## line plot

## Area

## Stacked area

## Streamchart

## time series

# Good for geograpical data

## Map

## Choropleth

## Hexbin map

## Cartogram

## Connection

## Bubble map

# Good for showing flow

## Chord diagram

## Network

## Sankey

## Arc diagram

## Edge Bundling

#


{% include links.md %}
