---
title: "types of plots"
teaching: 30
exercises: 15
questions:
- "What is JSON format?"

objectives:
- "Describe the JSON data format"
keypoints:
- "JSON is a popular data format for transferring data used by a great many Web based APIs"

source: Rmd
---

```{r, include=FALSE}
source("../bin/chunk-options.R")
knitr_fig_path("05-")
source("../bin/download_data.R")
## load the tidyverse
library(tidyverse)
```

# gode til fordelinger

## violin

```{r violin}
mpg$class = with(mpg, reorder(class, hwy, median))

mpg %>%
  ggplot( aes(x=class, y=hwy, fill=class)) + 
    geom_violin() +
    xlab("class") +
    theme_void() +
    theme(legend.position="none") +
    xlab("")
   



```


## density
```{r density, purl = FALSE, eval = TRUE, echo = FALSE}
ggplot(diamonds, aes(carat)) +geom_density(fill="#69b3a2", color="#e9ecef", alpha=0.8) +
  xlim(c(0,3))
```

## Histogram

```{r histogram, purl = FALSE, eval = TRUE, echo = FALSE}
data=data.frame(value=rnorm(100))

# basic histogram
ggplot(data, aes(x=value)) + 
  geom_histogram()
```


## Boxplot

```{r boxplot, purl = FALSE, eval = TRUE, echo = FALSE}
ggplot(mtcars, aes(x=as.factor(cyl), y=mpg)) + 
    geom_boxplot(fill="slateblue", alpha=0.2) + 
    xlab("cyl")
```


## Ridgeline

```{r ridgeline, purl = FALSE, eval = TRUE, echo = FALSE}
library(ggridges)
ggplot(lincoln_weather, aes(x = `Mean Temperature [F]`, y = `Month`, fill = ..x..)) +
  geom_density_ridges_gradient(scale = 3, rel_min_height = 0.01) +
  labs(title = 'Temperatures in Lincoln NE in 2016') +
  theme(
      legend.position="none",
      panel.spacing = unit(0.1, "lines"),
      strip.text.x = element_text(size = 8)
    )
```


# Gode til korrelationer

## scatter
Basic scatterplot
```{r scatter, purl = FALSE, eval = TRUE, echo = FALSE}
ggplot(iris, aes(x=Sepal.Length, y=Sepal.Width)) + 
    geom_point()
```

alternativ:
```{r, purl = FALSE, eval = TRUE, echo = FALSE}
ggplot(iris, aes(x=Sepal.Length, y=Sepal.Width)) + 
    geom_point(
        color="black",
        fill="#69b3a2",
        shape=22,
        alpha=0.5,
        size=6,
        stroke = 1
        ) 
```


## Heatmap

## Correlogram

## Bubble

## Connected scatter

## Density 2d

# Gode til ranking

## Barplot


```{r, purl = FALSE, eval = TRUE, echo = FALSE}
ggplot(mtcars, aes(x=as.factor(cyl), fill=as.factor(cyl) )) +  
  geom_bar( ) +
  scale_fill_manual(values = c("red", "green", "blue") ) +
  theme(legend.position="none")
```
If there is no natural order of the x-values, consider reordering 
by size:

```{r, purl = FALSE, eval = TRUE, echo = FALSE}
mtcars %>% 
  mutate(cyl = factor(cyl)) %>% 
  mutate(cyl = fct_infreq(cyl)) %>% 
ggplot(aes(x=cyl, fill=as.factor(cyl) )) +  
  geom_bar( ) +
  scale_fill_manual(values = c("red", "green", "blue") ) +
  theme(legend.position="none")
```


## Spider/radarplot

## Wordcloud

## Parallel

## Lollipop

```{r lollypop, purl = FALSE, eval = TRUE, echo = FALSE}
# Create data
data <- data.frame(
  x=LETTERS[1:26],
  y=abs(rnorm(26))
)

# Horizontal version
ggplot(data, aes(x=x, y=y)) +
  geom_segment( aes(x=x, xend=x, y=0, yend=y), color="skyblue") +
  geom_point( color="blue", size=4, alpha=0.6) +
  theme_light() +
  coord_flip() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank()
  )
```
Grundlæggende et barchart. 

## Cirkulær barplot

# gode til at vise dele af en helhed

## grupperet og stakket barplot

## Treemap

## Doughnut

Lige så meget skrammel som lagkagediagrammer.
```{r doughnut, purl = FALSE, eval = TRUE, echo = FALSE }
data <- data.frame(
  category=c("A", "B", "C"),
  count=c(10, 60, 30)
)
 
# Compute percentages
data$fraction = data$count / sum(data$count)

# Compute the cumulative percentages (top of each rectangle)
data$ymax = cumsum(data$fraction)

# Compute the bottom of each rectangle
data$ymin = c(0, head(data$ymax, n=-1))
 
# Make the plot
ggplot(data, aes(ymax=ymax, ymin=ymin, xmax=4, xmin=3, fill=category)) +
     geom_rect() +
     coord_polar(theta="y") + # Try to remove that to understand how the chart is built initially
     xlim(c(2, 4)) # Try to remove that to see how to make a pie chart
```


## Pie Chart

```{r pie-chart, purl = FALSE, eval = TRUE, echo = FALSE }
# Create Data
data <- data.frame(
  group=LETTERS[1:5],
  value=c(13,7,9,21,2)
)
# Compute the position of labels
data <- data %>% 
  arrange(desc(group)) %>%
  mutate(prop = value / sum(data$value) *100) %>%
  mutate(ypos = cumsum(prop)- 0.5*prop )

# Basic piechart
ggplot(data, aes(x="", y=prop, fill=group)) +
  geom_bar(stat="identity", width=1, color="white") +
  coord_polar("y", start=0) +
  theme_void() + 
  theme(legend.position="none") +
  
  geom_text(aes(y = ypos, label = group), color = "white", size=6) +
  scale_fill_brewer(palette="Set1")
```

Men pie-charts er noget skrammel


```{r}
library(patchwork)
data1 <- data.frame( name=letters[1:5], value=c(17,18,20,22,24) )
data2 <- data.frame( name=letters[1:5], value=c(20,18,21,20,20) )
data3 <- data.frame( name=letters[1:5], value=c(24,23,21,19,18) )
plot_pie <- function(data, vec){

ggplot(data, aes(x="name", y=value, fill=name)) +
  geom_bar(width = 1, stat = "identity") +
  coord_polar("y", start=0, direction = -1) +
  geom_text(aes(y = vec, label = rev(name), size=4, color=c( "white", rep("black", 4)))) +
  scale_color_manual(values=c("black", "white")) +
  theme(
    legend.position="none",
    plot.title = element_text(size=14),
    panel.grid = element_blank(),
    axis.text = element_blank(),
    legend.spacing=unit(0, "null")
  ) +
  xlab("") +
  ylab("")
  
}
a <- plot_pie(data1, c(10,35,55,75,93))
b <- plot_pie(data2, c(10,35,53,75,93))
c <- plot_pie(data3, c(10,29,50,75,93))
a + b + c
```

```{r}

plot_bar <- function(data){
  ggplot(data, aes(x=name, y=value, fill=name)) +
    geom_bar( stat = "identity") +
    scale_color_manual(values=c("black", "white")) +
    theme(
      legend.position="none",
      plot.title = element_text(size=14),
      panel.grid = element_blank(),
    ) +
    ylim(0,25) +
    xlab("") +
    ylab("")
}

# Make 3 barplots
a <- plot_bar(data1)
b <- plot_bar(data2)
c <- plot_bar(data3)

# Put them together with patchwork
a + b + c
```


## Dendrogram

## Circular packing

# Good for showing evolution

## line plot

## Area

## Stacked area

## Streamchart

## time series

# Good for geograpical data

## Map

## Choropleth

## Hexbin map

## Cartogram

## Connection

## Bubble map

# Good for showing flow

## Chord diagram

## Network

## Sankey

## Arc diagram

## Edge Bundling

#


{% include links.md %}
